---
title: "kmeans"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{kmeans}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# How to install the kmeans library 

```{r setup, eval = FALSE}
library(devtools)
devtools::install_github("zhaodyleo/Kmeans")
```

# One-dimension clustering

## Example 

The `kmeans_cluster` function in `kmeans` library could perform one dimension clustering
```{r}
library(kmeans)

# randomly generate training data
x1 <- runif(10000,0,1)
x2 <- runif(10000,2,3)
x3 <- runif(10000,4,5)
train_data <- c(x1,x2,x3)
train_label <- c(rep(0,10000),rep(1,10000),rep(2,10000))
training_result <- kmeans_cluster(train_data, 3, 2500, 1e-8)
# there is two output in training result 
# one is centre 
training_result[[1]]
# One is the predict label 
training_result[[2]][1:10]
```

We could use the `kmeans_cluster_predict` function to predict the label which new observations belongs. 
```{r}
X2 <- runif(1000,0,5)
kmeans_cluster_predict(X2,training_result[[1]])[1:10]
```

Sometimes, the label we got from the kmeans didn't match the actual label. Thus,we could use `kmeans_label_rematch` function to match the train label with the actual label. This function will also return the training accuracy. If we input the test data, and the actual test label. It will return the test accuracy and the rematched label lists.
```{r}
rematch_list <- kmeans_label_rematch(training_result[[2]],train_label)
rematch_list[[1]][1:10]
# training accuracy
rematch_list[[2]]
```
```{r}
test_label <- c(1,2,2,1,3,2,1)
test_actual<- c(2,0,0,2,1,0,2)
rematch_list_test <- kmeans_label_rematch(training_result[[2]],train_label,test_label)
# rematched training label
rematch_list_test[[1]][1:10]
# rematched test label
rematch_list_test[[2]]
# training accuracy
rematch_list_test[[3]]

# include the actual test label will give the test accuracy
rematch_list_test <- kmeans_label_rematch(training_result[[2]],train_label,test_label,test_actual)
rematch_list_test[[4]]
```

## Comparision with the kmeans function from `stats` library

Following code will test the accuracy of kmeans_cluster with the kmeans function from stats library
```{r}
x1 <- runif(100000,0,1)
x2 <- runif(100000,2,3)
x3 <- runif(100000,4,5)

index <- sample(1:300000,replace = FALSE)
train_data <- c(x1,x2,x3)[index]
train_label <- c(rep(0,10000),rep(1,10000),rep(2,10000))[index]
training_result <- kmeans_cluster(train_data, 3, 2500, 1e-8)
kmeans_stat_result <- kmeans(train_data,3)
kmeans_cluster_label <- kmeans_label_rematch(training_result[[2]],train_label)
kmeans_label <- kmeans_label_rematch(kmeans_stat_result$cluster,train_label)

all.equal(kmeans_cluster_label,kmeans_label)
```

Let's compare the efficiency between kmeans_cluster and kmeans from stats.
```{r}
library(bench)
bench::mark(kmeans_cluster(train_data, 3, 2500, 1e-8))
bench::mark(kmeans(train_data,3,iter.max = 2500))
```


